Modular Synthesis of Verified Verifiers of Computation with STV Algorithms

### 2019-03-14 paper/paper.tex
small changes

The goal of this paper is to advance formal methods in elections by demonstrating how STV vote counting can be verified in a modular, transparent, and practical feasible way over a minimal computational trust base:  we show how to synthesise efficient and provably correct verifiers for a large class of STV schemes in a modular and almost fully automated.

@@

We observe that every instance of STV performs the same \emph{actions} (e.g. count first preferences, transfer a surplus, eliminate a candidate, etc.) where the precise interpretation of action differs. We formalise these actions as rules that advance the state of the count towards a final result. Evidence for an instance of computation then becomes a sequence of rule applications applied to perform this computation. While generic STV abstracts the commonalities, each \emph{instance} of STV formally specifies the
pre- and postconditions for each rule as well as a boolean function that decides whether a rule has been applied correctly.

Executable evidence verifiers are constructed almost
automatically:  almost all of the ensuing proof obligations are discharged generically, and we fully  automate the subsequent translation into machine-level verified executable using CakeML.

@@

We are motivated to 
%engineer a trustworthy practical  framework to answer the 
tackle the following problem. How can one verify that any execution of any implementation $\mathcal{P}$ whose source code is secret of an arbitrary  known STV algorithm $\mathcal{A}$ correctly computes the end result according to  $\mathcal{A}$?

To begin 
%with addressing the problem, 
assume $\mathcal{P}$ is an implementation of an algorithm $\mathcal{A}$. We reasonably demand each execution of $\mathcal{P}$ on a given input $x$ consisting of ballots recorded in the election to output the winners $y$ and evidence $\omega$ as claim  for correctness of the execution.  
 Having evidence available for each execution of $\mathcal{P}$ facilitates checking correctness of the computation carried out \emph{independently} of (the source code of) $\mathcal{P}$. Therefore the original problem boils down to how one can verify the evidence of any \emph{instance of computation} with any algorithm $\mathcal{A}$.

The evidence as such must be enough informative to provide transparency of tallying and also allow  voters, or at least a large pool of scrutineers, to verify for themselves that the tally is authentically processed as per instructions of the counting scheme  so that winners truly reflect will of the voters. Therefore the question becomes what information is enough  for establishing transparency and verifiability of tallying  and what data structure for recording the information in the evidence we should choose.
\subsection{Characterisation of Evidence} 
A comprehensive analysis of different STV algorithms reveals existence of a common underlying abstract data type among STV algorithms as follows. 
%Each STV scheme operates on the following data.
\item Some ballots to deal with (assuming) recorded as cast by voters. Each ballot is a pair consisting of a list expressing the preference of a voter and a fractional transfer value,
\item a quota which is the least amount of votes needed to attract in order to be elected, 
\item some candidates competing in the election, 
\item some number of vacancies for which candidates compete,
\item \emph{discrete states of computation} which encapsulate necessary  pieces of information needed to transparently know how the tallying has progressed from the beginning to the end and to be able to independently verify its correctness provided access to the information is granted. 
There are two kinds of discrete states of computation. Some of them are final states where tallying has reached a halting state by finding the winners of the election. The other consist of non-final ones where bits and pieces of data are manipulated according to counting mechanism of the STV scheme used, to eventually obtain an end result. Each non-final state comprises seven pieces of data: 
 \item\textbf{uncounted ballots} which await being counted,
 \item\textbf{tally} for holding information about the amount of votes that each candidate has hitherto attracted %up to this state of computation,
 \item\textbf{pile} to know the ballots allocated to each candidate, 
 \item\textbf{backlog of elected} candidates who have exceeded the quota and their surplus votes awaits being transferred,
 \item\textbf{backlog of eliminated} ones whose votes awaits transfer,
 \item\textbf{elected candidates} who are the already elected ones,% whose surplus votes must be transferred,
 \item\textbf{continuing candidates} still continuing in the election. 
The non-final and final states specified above provide us with all information one needs. We therefore find a perfect characterisation for evidence. 
\begin{definition}\label{formalEv}
Suppose an STV algorithm $\mathcal{A}$ is given and $Q$, $s$ and $C$ are respectively the quota, number of vacancies and competing candidates in an electon whose counting algorithm is $\mathcal{A}$ . Also assume  $y$ is the output of an execution of $\mathcal{A}$ on an input $x$. By evidence $\hat{\omega}$ for the execution  we mean the quadruple $(Q,s,C,\Omega)$, where $\Omega$ is the chronological sequence of all states of this instance of computation visited from the initial state where $x$ is input to the final state where $y$ is output.   
\section{The Generic STV  Machine}
\label{sec:Machine}
\setlength{\abovedisplayskip}{0.5em}
\setlength{\belowdisplayskip}{0.5em}
We have already discussed an abstract data type whose values  formally represent evidence. However we do not yet possess an operational semantics for manipulating data to validate evidence.   On one hand, the semantics must be flexible enough to accommodate variations existing in the counting mechanism of STV algorithms so that we can produce verifiers  specifically operating for validation of evidence output by their associated scheme. On the other hand it must facilitate designing and developing the framework modularly while offering automation of synthesising process of verifiers.

In a secondary examination of the STV family, we also identify a general algorithmic mechanism for performing computation on evidence, which realises the \emph{commonalities} among the family members. This generic method comprises seven general actions%~(Figure~\ref{GenSteps})  
~(Definition~\ref{transitions}) corresponding to actions that tally officers take when counting votes.
\begin{definition}[Generic Counting Actions]\label{transitions}
 The following comprise the generic STV counting actions.
 %The generically appearing counting actions in the STV family are the followings. 
%\begin{figure}  
 \begin{itemize}
\item\textbf{count} for counting the uncounted ballots,
\item\textbf{elect} for electing all or some of the electable candidates, 
\item\textbf{transfer-elected} for distributing surplus votes of elected, %candidates, 
\item\textbf{eliminate} to exclude one or some candidates, 
\item\textbf{transfer-removed} to distribute votes of   the eliminated, %candidates, 
\item\textbf{elected-win} for terminating tallying whenever the vacancies are filled or there is no continuing candidate left, 
\item\textbf{hopeful-win} for terminating tallying whenever  the number of already elected and continuing candidates collectively does not exceed the initial vacancies. 
\end{itemize}
%\caption{Generic STV Counting Steps}
%\label{GenSteps}
%\end{figure}
\end{definition}
Each action is constituted of  preconditions on when and how the action comes into effect, and postconditions dictating what effect applying the action brings about by describing the new state to which computation proceeds. In fact, these conditions are the formal counterparts of the legal clauses in the textual specification of the counting scheme which tell tally officers how to proceed with tallying, and they are common between all flavours of STV.

For example a constant precondition for applying elimination  asserts that there must be no uncounted ballot left and another one declares that no candidate should be electable at the current state of counting. A postcondition of elimination is that once a candidate is excluded from counting, the list of continuing candidates is updated accordingly so that the excluded candidate no longer received votes. Conjunctions of such declarations comprise what an action is and what an application of that action means.

The underlying abstract data type and the universal algorithmic pattern found in STV characterise components of a finite state machine which we call \emph{generic STV}~(Definition\ref{STVMachine}). The set of machine states consist of the discrete states of computation $\mathcal{S}$ mentioned earlier. The seven generic actions given in Definition~\ref{transitions} form the transition labels $\mathcal{T}$ of the machine. Last, the everywhere present pre- and post-conditions which define (content of)  the actions, make a small-step operational semantics for the machine.  
Note that the states of computation recorded in all possible  evidence now become identical with machine states.

\begin{definition}[The generic STV machine]\label{STVMachine}
Let $\mathcal{S}$  and $\mathcal{T}$
be respectively the set of machine states and transition labels.  Also assume for $t\in\mathcal{T}$, $S_{t}$ $=$ $\bigwedge_{\atop{i\leq j_{t}}} \psi_{i}$ 
 where $\psi_{i}$ is the formal specification (in higher-order logic) of a pre- or post-condition of $t$. Then the \emph{generic STV} is the triple $M = \langle \mathcal{S}, \mathcal{T}, (S_t)_{t \in \mathcal{T}} \rangle$.  
\end{definition}
There are nonetheless variations in STV algorithms  which we  recognise  by separate \emph{instantiations} into the machine.  
\subsection{Instantiations of the Machine with Specific STV schemes}
Differences exist between counting algorithms of STV schemes.  For example in the STV used for electing the upper house representatives of the Victoria state of Australia (Victoria STV)~\cite{vic}, distributing votes of an eliminated candidate occurs step by step in order of the magnitude of the fractional values that those votes carry.  Consequently, a pre-condition for applying the action is to first assure that upon each instance of applying transfer-removed action (a) exactly those votes which have the same fractional value are transferred and (b) the transfer happens according to the magnitude order. In contrast, for example, the STV used in the lower house elections of Tasmania state of Australia (Tasmania STV) does not have such precondition simply because votes of an eliminated candidate are transferred in a single application of transfer-removed. There are also variations in postconditions. For example transfer-elected of the ACT STV~\cite{act} used in the lower house elections of the ACT state of Australia requires distributing only the last parcel of surplus votes received by an elected candidate. But Victoria STV specifies transferring all surplus votes of an elected candidate.

@@

\begin{definition}[verifier]\label{verifier}
Assume $\hat{\mathcal{A}}= \langle \mathcal{S}, \mathcal{T}, (S'_t)_{t \in \mathcal{T}} \rangle$ is an instantiation of the STV algorithm $\mathcal{A}$. A verifier for instances of computation of $\mathcal{A}$ is a function $\hat{\mathcal{V}}$ mapping from $\mathbb{Q}\times\mathbb{N}\times 2^{C}\times$ \textsf{List}$(\mathcal{S})$ to $\{0,1\}$ such that for any evidence $\hat{\omega} = (Q,s,C,\Omega)$ where $\Omega=\langle\Omega_{1},\dots,\Omega_{n}\rangle$, $\hat{\mathcal{V}}(\hat{\omega}) = 1$ if and only if the following holds:\\
$\forall i\in\{1,\dots,n-1\}.$ 
%\hspace*{0.2cm}$\forall \Omega_{i-1}$ $\Omega_{i}\in\mathcal{S}.$
%$\exists \Omega_{1}$ $\Omega_{2}\in$ \textsf{List} $(\mathcal{S}).$ $\Omega = \Omega_{1} + [\Omega_{i-1};\Omega_{i}] + \Omega_{2}$  
$\exists t\in\mathcal{T}.$  $(\Omega_{i-1},\Omega_{i})\in t$ $\wedge~\vdash S'_{t}[\Omega_{i-1},\Omega_{i}]$\\ 
where $\mathbb{Q}$ and $\mathbb{N}$ respectively represent the set of rational and natural numbers, $2^{C}$ is the set of all subsets of $C$, and \textsf{List}$(\mathcal{S})$ is the set of all possible lists of machine states. $\Omega_{i-1}$ is a pre-state and $\Omega_{i}$ is a post-state visited in the execution. The last line means that pre- and post-conditions  of applying $t$ to move from $\Omega_{i-1}$ to $\Omega_{i}$  are logically true. 
\end{definition} 
We shall next provide an example to concretely illustrate a piece of evidence and exemplify how STV schemes work. 
 \subsection{An Example of a Concrete Piece of Evidence}
\label{subsec:InstEv}
Figure~\ref{EvInst} depicts evidence of a small election whose counting algorithm is the ACT STV 
with three candidates $a$, $b$ and $c$, and initially recorded ballots containing
$b_1=([a,c],1/1)$, $b_2=([a,b,c],1/1)$, $b_3=([a,c,b],1/1)$,
$b_4=([b,a],1/1)$, $b_5=([c,b,a],1/1)$.  The first component of each ballot is a preference-ordered list of candidates, and the second is the fractional transfer value (initially set to $1/1$). The evidence consists of a \emph{header} that specifies the quota (computed according to the ACT STV instructions), the number of vacancies, and the list of competing candidates. 
The fourth line is the election result, and the remainder of the evidence are non-final states visited to compute this
outcome. Each non-final state shows uncounted ballots, tallies of candidates, piles of votes attracted, backlog of the elected, backlog of the eliminated, list of elected candidates, and list of continuing candidates.

First preferences for each candidate are computed. Ballots counted in favour of particular candidates are placed onto that candidate’s pile. In this election, $a$ is the first preference on $b_{1}$, $b_{2}$, and $b_{3}$, $b$ receives $b_{4}$, and $c$ receives $b_{5}$. Tallies are updated so that tally of $a$ becomes 3, and $b$
and $c$ each reach $1/1$. As $a$ exceeds the quota, he is elected and the fractional value of his surplus is updated according to the ACT scheme. Then $a$'s surplus is transferred to continuing candidates $b$ and $c$ and votes are counted according to next preferences shown on $a$'s surplus votes. Candidate $b$ and $c$ each attracts one vote. However  neither of them reaches to the quota. Hence the weakest one $b$ is eliminated. As the number of elected and continuing candidates is equal to 
%does not exceed 
the initial vacancies, the election terminates by declaring $a$ and $c$ winners.
\begin{small} 
\begin{figure}[b]
\begin{tabular}{c@{\hspace{2cm}}c} 
\AxiomC{\tiny 8/3} \noLine

@@

\section{The Framework Architecture}\label{sec:Arc} 
We progress through three macro-level phases to develop the framework. For the first phase we use HOL4 to formally specify, implement and verify the generic STV, its instantiations, verifiers, and every auxiliary assertions needed for these ends. The second step relies on the verified proof translation tool of CakeML to provably correctly translate implementations of the first phase into equivalent CakeML implementations. The last phase uses the ecosystem of CakeML and its verified compiler to generate machine executable evidence verifiers.

@@

\end{tikzpicture}
\caption{Architecture of the Framework, note that direction of arrow represents module dependencies}
\label{PicArc}
\end{figure}
\paragraph*{\textbf{Compilation}} 
Instantiations of CakeML's  compiler for generating machine executable verifiers happen in this module. For any deeply embedded $\mathcal{V}^{*}_{i}$, corresponding to $\hat{\mathcal{V}}_{dec}$ in \textbf{STV}$_{i}$, using proofs  established in \textbf{DeepSpec}, we instantiate the compiler to synthesise executable  verifier $\mathcal{V}_{i}$.  
\section{The Generic STV Formalised}\label{sec:GenCertVer}
We proceed through three steps, namely specification, implementation and verification of implementations in HOL4 to formally verify the machine and its components.
\paragraph*{Specification}  
Recall that we require minimising the trusted computing base  for computing with generated verifiers. The specification module consists of properties which we can use to verify that our implementations of  STV algorithms in HOL4 indeed match with the description of the counting methods used.  
 In light of Definition~\ref{STVMachine} these properties   actually form the semantics of the generic STV machine.   
\paragraph*{Implementation} Our final objective for creating the framework is to produce means for validating concrete evidences. To this end, we need decision procedures used for deciding whether or not a given evidence is valid. Therefore we define boolean-valued functions in the implementation sub-module to computationally realise logical declarations in the specification module. In particular, we implement decision procedures whose computational content  provably realises the specification of the machine semantics.  
\paragraph*{Verification} Here we formally prove that the implementations meet the expectations of their respective specifications. Therefore through verification we come to eliminate a trusted layer  required to lay in our framework.

We also have another motivation for developing the auxiliary module through this three-part process of specification, implementation and then verification. Familiarity, but no expertise, with our framework and its general purpose is needed for extending it to synthesise an executable verifier for one's desired STV scheme. On the one hand, an average user may lack enough skills in grasping functional programming style which we rely on for implementing verifiers in HOL4. However, the same user most probably has had exposure to formulations of mathematical properties  in first-order logic syntax. Including purely descriptive logical assertions as our specifications, proven equivalent to their implementations, facilitates the users with means to understand the functionality of components, modules and the system as a whole simply by inspecting the specifications instead of implementations.

@@

 \item HOL4 has well developed libraries comprised of verified assertions of operations on list structure. By structuring the data on lists, we facilitate us and users  with exploitation of  already verified tools to formalise the framework and avoid inventing the unnecessary from the scratch. 
 \item HOL4 also has well-developed tactics for discharging proof obligations on assertions involving list structure and operations on it. We therefore come to provide us with means for ease of verification of the formalised assertions.
 \item Understanding the data structure used in the  framework implementation is critical for ease of  its usability and extensibility by third parties.  The type $\mathit{judgement}$ closely models concrete evidences such as the one in figure~\ref{EvInst}. Therefore one can sensibly perceive the  abstraction step taken for modelling  concrete data which in turn enhances understandability of the framework and its mechanism. 

@@

 Recall that objects whose type is $\mathit{piles}$ serve as containers for recording the ballots allocated to each candidate. Our choice for type of piles allows us to store the ballots received by each candidate in chunks of lists rather than one single list containing all of them.  We have two reasons for designing the type of piles as it stands instead of $\mathit{(cand\times(ballots~list))~list}$;  
\paragraph*{(a)} Some STV schemes such as the lower house ACT and Tasmania STV, employ a notion called ''last parcel''. In short the last parcel of a candidate is the collection of those ballots received by the candidate at the last round of application of the count transition which made the tally of the candidate reach or exceed the quota and therefore be elected.  Only the last parcel, instead of all, of an elected candidate is distributed, according to next preferences shown on the ballots in the last parcel, to the continuing candidates.

@@

One can already proceed to synthesise an executable verifier from the machine. Such a verifier can correctly decide if given evidence $\omega$ claimed to have been produced by an algorithm whose counting scheme is STV, instead of e.g. First-Past-The-Post voting~\cite{DFar} where voters choose whom they prefer most and the candidate with the most amount of votes received is elected.  
However we wish to generate verifiers that can recognise and validate according to which specific STV algorithm the evidence $\omega$ has been output. Hence we need to enrich the computational content of the machine semantics with more  pre- and postconditions which are particular to individual STV schemes. We refer to this enrichment process as instantiation of the machine and discuss it further below.

@@

\item[$\bullet_{3}$] distribution of the surplus of an elected candidate proceeds in one single step.
\item[$\bullet_{4}$] surplus of elected candidates is distributed one at a time beginning with those who are elected earlier. 
\item[$\bullet_{5}$] pile of the candidate whose surplus is transferred is emptied in the post-state of transfer-elected.

@@

\item[$\bullet_{7}$] pile of any candidate other than the one whose surplus is transferred at this stage remains the same.

>>>>

The goal of this paper is to advance formal methods in elections by
demonstrating how STV vote counting can be verified in a modular,
transparent, and practical feasible way over a minimal computational
trust base:  we show how to synthesise efficient and provably
correct verifiers for a large class of STV schemes in a modular and
almost fully automated way.

@@

We observe that every instance of STV performs the same
\emph{actions} (e.g. count first preferences, transfer a surplus,
eliminate a candidate, etc.) where the precise interpretation of
actions differs. We formalise these actions as rules that advance the state of the count towards a final result. Evidence for an instance of computation then becomes a sequence of rule applications applied to perform this computation. While generic STV abstracts the commonalities, each \emph{instance} of STV formally specifies the
pre- and postconditions for each rule as well as a boolean function that decides whether a rule has been applied correctly.

Executable evidence verifiers are constructed almost
automatically:  nearly all of the ensuing proof obligations are discharged generically, and we fully  automate the subsequent translation into machine-level verified executable using CakeML.

@@

Our problem is the following: 
%engineer a trustworthy practical  framework to answer the 
how can we verify that an execution of an implementation $\mathcal{P}$ 
%whose source code is secret 
of an arbitrary  known STV algorithm $\mathcal{A}$ correctly
computes the end result according to  $\mathcal{A}$, even if
$\mathcal{P}$'s source code is secret?

To begin 
%with addressing the problem, 
assume $\mathcal{P}$ is an implementation of an algorithm
$\mathcal{A}$. We require that each execution of $\mathcal{P}$ on a
given input $x$ (of ballots cast in an election) to output both
winners $y$ and evidence $\omega$ as proof of correctness of the execution.  
 Having evidence available for each execution of $\mathcal{P}$
 facilitates checking correctness of the computation carried out
 \emph{independently} of (the source code of) $\mathcal{P}$.
 Therefore the original problem boils down verifying (the
 correctness of) evidence produced by an an
 \emph{instance of computation} of an algorithm $\mathcal{A}$.

The evidence as such must be enough informative to provide transparency of tallying and also allow  voters, or at least a large pool of scrutineers, to verify for themselves that the tally is authentically processed as per instructions of the counting scheme  so that winners truly reflect will of the voters. Therefore the question becomes what information is enough  for establishing transparency and verifiability of tallying  and what data structure for recording the information in the evidence we should choose.
\subsection{Characterisation of Evidence} 
A comprehensive analysis of different STV algorithms reveals existence of a common underlying abstract data type among STV algorithms as follows. 
%Each STV scheme operates on the following data.
\begin{itemize}
\item Some ballots to deal with (assuming) recorded as cast by voters. Each ballot is a pair consisting of a list expressing the preference of a voter and a fractional transfer value,
\item a quota which is the least amount of votes needed to be elected, 
\item the candidates competing in the election, 
\item the number of vacancies for which candidates compete,
\item the \emph{discrete states of computation} which encapsulate necessary  pieces of information needed to transparently know how the tallying has progressed from beginning to end and to be able to independently verify its correctness provided access to the information is granted. 
\end{itemize}
There are two kinds of discrete states of computation. Some of them are final states where tallying has reached a halting state by finding the winners of the election. The other consist of non-final ones where bits and pieces of data are manipulated according to counting mechanism of the STV scheme used, to eventually obtain an end result. Each non-final state comprises seven pieces of data: 
\begin{itemize}
 \item\textbf{uncounted ballots} which await being counted,
 \item\textbf{tally} for holding information about the amount of votes that each candidate has hitherto attracted %up to this state of computation,
 \item\textbf{pile} to know the ballots allocated to each candidate, 
 \item\textbf{backlog of elected} candidates who have exceeded the quota and their surplus votes awaits being transferred,
 \item\textbf{backlog of eliminated} ones whose votes awaits transfer,
 \item\textbf{elected candidates} who are the already elected ones,% whose surplus votes must be transferred,
 \item\textbf{continuing candidates} still continuing in the election. 
 \end{itemize}
The non-final and final states specified above provide us with all information one needs. We therefore find a perfect characterisation for evidence. 
\begin{definition}\label{formalEv}
Suppose an STV algorithm $\mathcal{A}$ is given and $Q$, $s$ and $C$
are respectively the quota, number of vacancies and competing
candidates in an election whose counting algorithm is $\mathcal{A}$. Also assume  $y$ is the output of an execution of $\mathcal{A}$ on an input $x$. By evidence $\hat{\omega}$ for the execution  we mean the quadruple $(Q,s,C,\Omega)$, where $\Omega$ is the chronological sequence of all states of this instance of computation visited from the initial state where $x$ is input to the final state where $y$ is output.   
\end{definition}
\section{The Generic STV  Machine}
\label{sec:Machine}
\setlength{\abovedisplayskip}{0.5em}
\setlength{\belowdisplayskip}{0.5em}
We have already discussed an abstract data type whose values
formally represent evidence. However we do not yet possess an
operational semantics for manipulating data to validate evidence.
On one hand, the semantics must be flexible enough to accommodate
variations existing in the counting mechanism of STV algorithms so
that we can produce verifiers  specifically operating for validation
of evidence output by their associated scheme. On the other hand it
must facilitate designing and developing the framework modularly
while offering automation for synthesising verifiers.

We also identify a general algorithmic mechanism for performing
computation on evidence, which realises the \emph{commonalities}
among the members of the STV family. This generic method comprises seven general actions%~(Figure~\ref{GenSteps})  
~(Definition~\ref{transitions}) corresponding to actions that tally officers take when counting votes.
\begin{definition}[Generic Counting Actions]\label{transitions}
 The following comprise the generic STV counting actions.
 %The generically appearing counting actions in the STV family are the followings. 
%\begin{figure}  
 \begin{itemize}
\item\textbf{count} for counting the uncounted ballots,
\item\textbf{elect} for electing all or some of the electable candidates, 
\item\textbf{transfer-elected} for distributing surplus votes of elected, %candidates, 
\item\textbf{eliminate} to exclude one or some candidates, 
\item\textbf{transfer-removed} to distribute votes of   the eliminated, %candidates, 
\item\textbf{elected-win} for terminating tallying whenever the vacancies are filled or there is no continuing candidate left, 
\item\textbf{hopeful-win} for terminating tallying whenever  the number of already elected and continuing candidates collectively does not exceed the initial vacancies. 
\end{itemize}
%\caption{Generic STV Counting Steps}
%\label{GenSteps}
%\end{figure}
\end{definition}
Each action consists of  preconditions on when and how the action
comes into effect, and postconditions dictating what effect applying
the action brings about by describing the new state to which the
computation proceeds. In fact, these conditions are the formal
counterparts of the legal clauses in the textual specification of
the counting scheme which inform tally officers how to proceed with tallying, and they are common between all flavours of STV.

For example, a constant precondition for applying elimination
asserts that there must be no uncounted ballot left and that no
candidate should be electable at the current state of counting. A
postcondition of elimination is that once a candidate is excluded
from counting, the list of continuing candidates is updated
accordingly so that the excluded candidate no longer received votes.
Conjunctions of such declarations define what an action is and what an application of that action means.

The underlying abstract data type and the universal algorithmic
pattern found in STV characterise components of a finite state
machine which we call \emph{generic
STV}~(Definition\ref{STVMachine}). The set of machine states consist
of the discrete states of computation $\mathcal{S}$ mentioned
earlier. The seven generic actions given in
Definition~\ref{transitions} form the transition labels
$\mathcal{T}$ of the machine. Last, the everywhere present pre- and
post-conditions which define (content of)  the actions, and
constitute a small-step operational semantics for the machine.  
Note that the states of computation recorded as evidence now become identical with machine states.

\begin{definition}[The generic STV machine]\label{STVMachine}
Let $\mathcal{S}$  and $\mathcal{T}$
be respectively the set of machine states and transition labels.  Also assume for $t\in\mathcal{T}$, $S_{t}$ $=$ $\bigwedge_{\atop{i\leq j_{t}}} \psi_{i}$ 
 where $\psi_{i}$ is the formal specification (in higher-order logic) of a pre- or post-condition of $t$. Then the \emph{generic STV} is the triple $M = \langle \mathcal{S}, \mathcal{T}, (S_t)_{t \in \mathcal{T}} \rangle$.  
\end{definition}
There are nonetheless variations in STV algorithms  which we  recognise  by separate \emph{instantiations} into the machine.  
\subsection{Instantiations of the Machine with Specific STV schemes}
\noindent
Differences exist between counting algorithms of STV schemes.  For
example in the STV used for electing the upper house representatives
of the Australian State of Victoria (Victoria STV)~\cite{vic},
distributing votes of an eliminated candidate occurs step by step in
order of the magnitude of the fractional values that those votes
carry.  Consequently, a pre-condition for applying the action is to
first assure that upon each instance of applying transfer-removed
action (a) exactly those votes which have the same fractional value
are transferred and (b) the transfer happens according to the
magnitude order. In contrast, for example, the STV used in the lower
house elections of the Australian state of Tasmania (Tasmania STV)
does not have such precondition simply because votes of an
eliminated candidate are transferred in a single application of
transfer-removed. There are also variations in postconditions. For
example transfer-elected of the ACT STV~\cite{act} used in the lower
house elections of the Australian Capital Territory  requires distributing only the last parcel of surplus votes received by an elected candidate. But Victoria STV specifies transferring all surplus votes of an elected candidate.

@@

\begin{definition}[verifier]\label{verifier}
Assume $\hat{\mathcal{A}}= \langle \mathcal{S}, \mathcal{T},
(S'_t)_{t \in \mathcal{T}} \rangle$ is an instantiation of the STV
algorithm $\mathcal{A}$. A \emph{verifier} for instances of
computation of $\mathcal{A}$ is a function $\hat{\mathcal{V}}$
mapping $\mathbb{Q}\times\mathbb{N}\times 2^{C}\times$
\textsf{List}$(\mathcal{S})$ to $\{0,1\}$ such that for any putative evidence $\hat{\omega} = (Q,s,C,\Omega)$ where $\Omega=\langle\Omega_{1},\dots,\Omega_{n}\rangle$, $\hat{\mathcal{V}}(\hat{\omega}) = 1$ if and only if the following holds:\\
$\forall i\in\{1,\dots,n-1\}.$ 
%\hspace*{0.2cm}$\forall \Omega_{i-1}$ $\Omega_{i}\in\mathcal{S}.$
%$\exists \Omega_{1}$ $\Omega_{2}\in$ \textsf{List} $(\mathcal{S}).$ $\Omega = \Omega_{1} + [\Omega_{i-1};\Omega_{i}] + \Omega_{2}$  
$\exists t\in\mathcal{T}.$  $(\Omega_{i-1},\Omega_{i})\in t$ $\wedge~\vdash S'_{t}[\Omega_{i-1},\Omega_{i}]$\\ 
where $\mathbb{Q}$ and $\mathbb{N}$ respectively represent the set
of rational and natural numbers, $2^{C}$ is the set of all subsets
of $C$, and \textsf{List}$(\mathcal{S})$ is the set of all possible
lists of machine states. The state $\Omega_{i-1}$ is a pre-state and $\Omega_{i}$ is a post-state visited in the execution. The last line means that pre- and post-conditions  of applying $t$ to move from $\Omega_{i-1}$ to $\Omega_{i}$  are logically true. 
\end{definition} 
We shall next provide an example to concretely illustrate a piece of evidence and exemplify how STV schemes work. 
 \subsection{An Example of a Concrete Piece of Evidence}
\label{subsec:InstEv}
\noindent
Figure~\ref{EvInst} shows evidence of a small election whose counting algorithm is the ACT STV 
with three candidates $a$, $b$ and $c$, and initially recorded ballots containing
$b_1=([a,c],1/1)$, $b_2=([a,b,c],1/1)$, $b_3=([a,c,b],1/1)$,
$b_4=([b,a],1/1)$, $b_5=([c,b,a],1/1)$.  The first component of each ballot is a preference-ordered list of candidates, and the second is the fractional transfer value (initially set to $1/1$). The evidence consists of a \emph{header} that specifies the quota (computed according to the ACT STV instructions), the number of vacancies, and the list of competing candidates. 
The fourth line is the election result, and the remainder of the evidence are non-final states visited to compute this
outcome. Each non-final state shows uncounted ballots, tallies of candidates, piles of votes attracted, backlog of the elected, backlog of the eliminated, list of elected candidates, and list of continuing candidates.

First preferences for each candidate are computed. Ballots counted in favour of particular candidates are placed onto that candidate’s pile. In this election, $a$ is the first preference on $b_{1}$, $b_{2}$, and $b_{3}$, $b$ receives $b_{4}$, and $c$ receives $b_{5}$. Tallies are updated so that tally of $a$ becomes 3, and $b$
and $c$ each reach $1/1$. As $a$ exceeds the quota, (s)he is elected and the fractional value of his surplus is updated according to the ACT scheme. Then $a$'s surplus is transferred to continuing candidates $b$ and $c$ and votes are counted according to next preferences shown on $a$'s surplus votes. Candidate $b$ and $c$ each attract one vote. However  neither of them reaches to the quota. Hence the weakest one $b$ is eliminated. As the number of elected and continuing candidates is equal to 
%does not exceed 
the initial vacancies, the election terminates by declaring $a$ and
$c$ to be winners.
\begin{small} 
\begin{figure}[b]
\begin{tabular}{c@{\hspace{2cm}}c} 
\AxiomC{\tiny 8/3} \noLine

@@

\section{The Framework Architecture}\label{sec:Arc} 
\noindent
We progress through three macro-level phases to develop the framework. For the first phase we use HOL4 to formally specify, implement and verify the generic STV, its instantiations, verifiers, and every auxiliary assertions needed for these ends. The second step relies on the verified proof translation tool of CakeML to provably correctly translate implementations of the first phase into equivalent CakeML implementations. The last phase uses the ecosystem of CakeML and its verified compiler to generate machine executable evidence verifiers.

@@

\end{tikzpicture}
\caption{Architecture of the Framework, note that direction of arrow represents module dependencies}
\label{PicArc}
\end{figure}
\paragraph*{\textbf{Compilation}} 
Instantiations of CakeML's  compiler for generating machine executable verifiers happen in this module. For any deeply embedded $\mathcal{V}^{*}_{i}$, corresponding to $\hat{\mathcal{V}}_{dec}$ in \textbf{STV}$_{i}$, using proofs  established in \textbf{DeepSpec}, we instantiate the compiler to synthesise executable  verifier $\mathcal{V}_{i}$.  
\section{Generic STV Formalised}\label{sec:GenCertVer}
We proceed through three steps, namely specification, implementation and verification of implementations in HOL4 to formally verify the machine and its components.
\paragraph*{Specification}  
Recall that we require minimising the trusted computing base  for computing with generated verifiers. The specification module consists of properties which we can use to verify that our implementations of  STV algorithms in HOL4 indeed match with the description of the counting methods used.  
 In light of Definition~\ref{STVMachine} these properties   actually form the semantics of the generic STV machine.   
\paragraph*{Implementation} Our final objective for creating the
framework is to produce means for validating concrete items of evidence. To this end, we need decision procedures used for deciding whether or not a given evidence is valid. Therefore we define boolean-valued functions in the implementation sub-module to computationally realise logical declarations in the specification module. In particular, we implement decision procedures whose computational content  provably realises the specification of the machine semantics.  
\paragraph*{Verification} Here we formally prove that the implementations meet the expectations of their respective specifications. Therefore through verification we come to eliminate a trusted layer  required to lay in our framework.

We also have another motivation for developing the auxiliary module through this three-part process of specification, implementation and then verification. Familiarity, but no expertise, with our framework and its general purpose is needed for extending it to synthesise an executable verifier for one's desired STV scheme. On the one hand, an average user may lack skills in functional programming which we rely on for implementing verifiers in HOL4. However, the same user most probably has had exposure to formulations of mathematical properties  in first-order logic syntax. Including purely descriptive logical assertions as our specifications, proven equivalent to their implementations, facilitates the users with means to understand the functionality of components, modules and the system as a whole simply by inspecting the specifications instead of implementations.

@@

 \item HOL4 has well developed libraries containing verified
 assertions of operations on lists. By using lists to represent the
 data, both us and users of this framework can 
exploit the (already
 verified) tools to formalise the framework and avoid re-proving
 assertions from scratch.
 \item HOL4 also has well-developed
 tactics for discharging proof obligations on assertions involving
 list structure and operations on it. This significantly eases the
 process of verification. 
 \item Understanding the data structure used in the  framework implementation is critical for ease of  its usability and extensibility by third parties.  The type $\mathit{judgement}$ closely models concrete evidence such as the one in Figure~\ref{EvInst}. Therefore one can sensibly perceive the  abstraction step taken for modelling  concrete data which in turn enhances understandability of the framework and its mechanism. 

@@

 Recall that objects whose type is $\mathit{piles}$ serve as
 containers for recording the ballots allocated to each candidate.
 Our choice for the  type of piles allows us to store the ballots received by each candidate in chunks of lists rather than one single list containing all of them.  We have two reasons for designing the type of piles as it stands instead of $\mathit{(cand\times(ballots~list))~list}$;  
\paragraph*{(a)} Some STV schemes such as the lower house ACT and
Tasmania STV, employ a notion called ''last parcel''. In short, the
last parcel of a candidate is the collection of those ballots
received by the candidate at the last round of application of the
count transition which caused the tally of the candidate to reach or
exceed the quota and therefore be elected.  Only the last parcel,
instead of all ballots received, of an elected candidate is distributed, according to next preferences shown on the ballots in the last parcel, to the continuing candidates.

@@

One can already proceed to synthesise an executable verifier from
the underspecified generic machine. Such a verifier can correctly decide if given evidence
$\omega$ claimed to have been produced by an algorithm of the STV
family,  instead of e.g. First-Past-The-Post
voting~\cite{DFar} where voters choose whom they prefer most and the
candidate with the most amount of votes received is elected.
However we wish to generate verifiers that can recognise and
validate according to specific instances of STV. 
$\omega$ has been output. Hence we need to enrich the computational
content of the machine semantics with more  pre- and postconditions
which are particular to individual STV schemes. We refer to this
enrichment process as instantiation of the machine and discuss it
further below.

@@

\item[$\bullet_{3}$] the distribution of the surplus of an elected candidate proceeds in one single step.
\item[$\bullet_{4}$] the surplus of elected candidates is distributed one at a time beginning with those who are elected earlier. 
\item[$\bullet_{5}$] the pile of the candidate whose surplus is transferred is emptied in the post-state of transfer-elected.

@@

\item[$\bullet_{7}$] the pile of any candidate other than the one whose surplus is transferred at this stage remains the same.
